Trail Saving on Backtrack

### 2020-05-05 paper.tex
minor mods

\begin{document}
\maketitle
\begin{abstract}
    A CDCL \sat solver can backtrack a large distance when it learns a
    new clause, e.g, when the new learnt clause is a unit clause the
    solver has to backtrack to level zero. When the length of the
    backtrack is large, the solver can end up reproducing many of the
    same decisions and propagations when it redescends the search
    tree. Motivated by this potential redundancy, partial backtracking
    (chronological backtracking) has been proposed. This technique
    allows the solver to reduce the length of its backtrack, e.g.,
    allowing the solver to backtrack to the previous level after a
    conflict is found. However, this technique has its shortcomings
    and only tends to work when limits are placed on its
    application. In this paper we present a new trail saving technique
    that makes a copy of the part of the trail that was backtracked
    over. This saved copy can then be used to improve the efficiency
    of the solver's subsequent redescent. Although our technique does
    not save as much work as chronological backtracking, it also does
    not have chronological backtracking's shortcomings. Furthermore,
    the saved trail also provides the solver with the ability to look
    ahead along the previous trail. This can be exploited to improve
    its efficiency. Our empirical results show that our technique
    often yields superior performance over chronological backtracking,
    and that is is able to improve the performance of state-of-the-art
    solvers.

@@

Each literal $\ell\in\trail$ has a decision level $\dlevel{\ell}$
which is equal to the number of decision literals appearing on the trail
up to and including $\ell$; hence, $\dlevel{d}=1$ for the first
decision literal $d\in\trail$. The set of literals on $\trail$ that
have the same decision level forms a \emph{contiguous}
subsequence\footnote{Our approach uses standard trails in which the
  decision levels are contiguous. Chronological backtracking
  \cite{DBLP:conf/lpar/JiangZ13,DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}
  generates trails with non-contiguous decision levels} that starts
with a decision literal $d_i$ and ends just before the next decision
literal $d_{i+1}$. We will often need to refer to different decision
level subsequences of $\trail$. Hence, we let $\trail[[i]]$ denote the
subsequence of literals at decision level $i$; and let
$\trail[[\range{i}{j}]]$ denote the subsequence of literals at
decision levels $k$ for $i\leq k\leq j$. 
\begin{defn}
    A clause $C$ has \textbf{been made unit by $\trail$ implying $l$}
    when
    $l\in C \land \bigl(\forall x\in C. x\neq l \implies \lnot x \in
    \trail\bigr)$. That is, all literals in $C$ except $l$ must have
    been falsified by $\trail$
\end{defn}

@@

This means that the decision levels on the trail are no longer
contiguous, as $\deepestLit$ has a different level than the other
literals at level $j$ (if $j\neq \btL$). This change has a number of
consequences for the \sat solver's operation, all of which were
described in the original paper \cite{DBLP:conf/lpar/JiangZ13}. In
\cite{DBLP:conf/sat/NadelR18} some more practical implementation
methods were developed to overcome this issues, and in
\cite{DBLP:conf/sat/MohleB19} a formal framework for \cbt was
developed under which the changes required to the \sat solver could be
shown to be sound. We refer the reader to the cited papers for
more details.

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT
and Cadical, both of which have finished at or near the top of SAT
competitions for the past several years
\cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each
solver on the 800 total benchmark instances used in the main tracks of
the 2018 and 2019 SAT competitions. The experiments were executed on a
cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB
memory limits for each instance. We chose not to output or verify the
proofs generated by any of the solvers. The Par-2 scores obtained and
total instances solved by each solver are reported in
Figures~\ref{fig:cadicalNew}, \ref{fig:cadicalOld}, and
\ref{fig:maple}. We also show the cactus plots of the new version of
cadical in Figure \ref{fig:cactus_cadical} and MapleLCMDist in Figure
\ref{fig:cactus_maple}.

>>>>

\begin{document}
\maketitle
\begin{abstract}
    A CDCL \sat solver can backtrack a large distance when it learns a
    new clause, e.g, when the new learnt clause is a unit clause the
    solver has to backtrack to level zero. When the length of the
    backtrack is large, the solver can end up reproducing many of the
    same decisions and propagations when it redescends the search
    tree. Motivated by this potential redundancy, partial backtracking
    (chronological backtracking) has been proposed. This technique
    allows the solver to reduce the length of its backtrack, e.g.,
    allowing the solver to backtrack to the previous level after a
    conflict is found. However, this technique has its shortcomings
    and only tends to work when limits are placed on its
    application. In this paper we present a new trail saving technique
    that makes a copy of the part of the trail that was backtracked
    over. This saved copy can then be used to improve the efficiency
    of the solver's subsequent redescent. Although our technique does
    not save as much work as chronological backtracking, it also does
    not have chronological backtracking's shortcomings. Furthermore,
    the saved trail also provides the solver with the ability to look
    ahead along the previous trail. This can be exploited to improve
    its efficiency. Our empirical results show that our technique
    often yields superior performance over chronological backtracking,
    and that is able to improve the performance of state-of-the-art
    solvers.

@@

Each literal $\ell\in\trail$ has a decision level $\dlevel{\ell}$
which is equal to the number of decision literals appearing on the trail
up to and including $\ell$ ; hence, $\dlevel{d}=1$ for the first
decision literal $d\in\trail$. The set of literals on $\trail$ that
have the same decision level forms a \emph{contiguous}
subsequence\footnote{Our approach uses standard trails in which the
  decision levels are contiguous. Chronological backtracking
  \cite{DBLP:conf/lpar/JiangZ13,DBLP:conf/sat/NadelR18,DBLP:conf/sat/MohleB19}
  generates trails with non-contiguous decision levels} that starts
with a decision literal $d_i$ and ends just before the next decision
literal $d_{i+1}$. We will often need to refer to different decision
level subsequences of $\trail$. Hence, we let $\trail[[i]]$ denote the
subsequence of literals at decision level $i$; and let
$\trail[[\range{i}{j}]]$ denote the subsequence of literals at
decision levels $k$ for $i\leq k\leq j$. 
\begin{defn}
    A clause $C$ has \textbf{been made unit by $\trail$ implying $l$}
    when
    $l\in C \land \bigl(\forall x\in C. x\neq l \implies \lnot x \in
    \trail\bigr)$. That is, all literals in $C$ except $l$ must have
    been falsified by $\trail$
\end{defn}

@@

This means that the decision levels on the trail are no longer
contiguous, as $\deepestLit$ has a different level than the other
literals at level $j$ (if $j\neq \btL$). This change has a number of
consequences for the \sat solver's operation, all of which were
described in the original paper \cite{DBLP:conf/lpar/JiangZ13}. In
\cite{DBLP:conf/sat/NadelR18} some more practical implementation
methods were developed to overcome these issues, and in
\cite{DBLP:conf/sat/MohleB19} a formal framework for \cbt was
developed under which the changes required to the \sat solver could be
shown to be sound. We refer the reader to the cited papers for
more details.

@@

\section{Experiments and Results}
We implemented our techniques in two different SAT solvers, MapleSAT
and Cadical, both of which have finished at or near the top of SAT
competitions for the past several years
\cite{Heule2018ProceedingsOS,Heule2019ProceedingsOS}. We then ran each
solver on the 800 total benchmark instances used in the main tracks of
the 2018 SAT Competition and 2019 SAT Race. The experiments were executed on a
cluster of 2.7 GHz Intel cores with 5000 seconds CPU time and 7 GB
memory limits for each instance. We chose not to output or verify the
proofs generated by any of the solvers. The Par-2 scores obtained and
total instances solved by each solver are reported in
Figures~\ref{fig:cadicalNew}, \ref{fig:cadicalOld}, and
\ref{fig:maple}. We also show the cactus plots of the new version of
cadical in Figure \ref{fig:cactus_cadical} and MapleLCMDist in Figure
\ref{fig:cactus_maple}.
